---
title: "eindproject"
author: "Cynthia de Nijs"
date: "20 mei 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(caret)
library(corrplot)
library(rpart)
library(rpart)
library(randomForest)
library(parallel)
library(doParallel)
library(ggplot2)
library(lattice)
```

### project course 8

class-variabele voorspellen
schrijf een rapport over hoe je het model hebt gebouwd
hoe je cross validation hebt gebruikt
wat je denkt dat de out of sample error is
waarom heb je de keuzes gemaakt die je hebt gemaakt.

Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

What you should submit

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

Peer Review Portion

Your submission for the Peer Review portion should consist of a link to a Github repo with your R markdown and compiled HTML file describing your analysis. Please constrain the text of the writeup to < 2000 words and the number of figures to be less than 5. It will make it easier for the graders if you submit a repo with a gh-pages branch so the HTML page can be viewed online (and you always want to make it easy on graders :-).

Course Project Prediction Quiz Portion

Apply your machine learning algorithm to the 20 test cases available in the test data above and submit your predictions in appropriate format to the Course Project Prediction Quiz for automated grading.

Reproducibility

Due to security concerns with the exchange of R code, your code will not be run during the evaluation by your classmates. Please be sure that if they download the repo, they will be able to view the compiled HTML version of your analysis.


# read the data and replace "NA", "#DIV/0!", "" by NA (160 columns)

```{r}
training <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"), na.strings=c("NA","#DIV/0!", ""))
testing <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"),  na.strings=c("NA","#DIV/0!", ""))
```

# cleaning data

remove NA-columns (60 columns left)
remove columns that will not explain the outcome variable: X, user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvdt_timestamp, new_window, num_window (53 columns left)

```{r}

training <- training[,colSums(is.na(training))==0]
testing <- testing[,colSums(is.na(testing))==0]

training <- training[,-c(1:7)]
testing <- testing[,-c(1:7)]

```

# split the data

 I split the trainingsset in a testset and a trainingset (train). The dataset isn't big and isn't small so the trainingsset contains 60% of the data and the testset 40%.

```{r}

set.seed(123)
inTrain <- createDataPartition(y=training$classe, p = 0.60, list=FALSE)
train <- training[inTrain,]
test <- training[-inTrain,]
dim(train)
dim(test)

```

# explore the data

 balance in outcomes? yes!

```{r}
table(train$classe)

```


 near zero variable; there are no near zero var. left

```{r}
nzv <- nearZeroVar(train, saveMetrics=TRUE)
nzv
```
how about correlations/multicollinearity? : a lot of blue -> multicollinearity

```{r}

corrplot(cor(train[,-53]), method="color")

```
let's have a look at the first 13 rows
```{r}

corrplot(cor(train[,c(1:13)]), method="color")

```


# make a model

classification model: Tree. this model is not accurate (was aspected because of multicollinearity)

```{r}
modFitT <- train(classe ~ ., method = "rpart", data=train)
finModT <- modFitT$finalModel
print(modFitT)

```

# classification model: random forest

let's try a random forest-model. This model can handle multicollinearity.

short discription:
the random forest-model creates a lot of trees. Each tree is build on a different sample of the data (bootstrapping). A each node of the tree is using a random set of m variables that may contribute to the split. The trees are voted to predict an outcome.


Because it's very slow we use clusters.

The chosen resampling method is cross validation (10-fold).

cv 5-fold: accuracy 0,99 
almost the same as 10-fold. for k=10 the bias-variance balance is a little bit beter.
also cv 30-fold has almost the same accuracy
The model seems to be stable. I use k=10.

```{r}
set.seed(123)
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)


fitControl <- trainControl(method = "cv",
                           number = 10,
                           allowParallel = TRUE)

modFitRf <- train(classe ~ ., method = "rf", data=train, trControl=fitControl)

stopCluster(cluster)
registerDoSEQ()

```

confusionmatrix:

```{r}
finModRf <- modFitRf$finalModel
finModRf
```

rf performs good and is accurate: 0,9907

```{r}
print(modFitRf)

``` 

#out of sample error
the accuracy of the testset is 0,9916 (see the statistics)
the out of sample error is 1-0,9916 =0,0084
the accuracy of the testset is (higer than)almost) the same as the accuracy of the train-set
```{r}
pred <- predict(modFitRf, test)
confusionMatrix(pred, test$classe)

```

# the most important variables

```{r}
varImp(modFitRf)

```


# now we predict on the testingset

```{r}
predT <- predict(modFitRf, testing)
predT

```